---
title: "Analysis of shooter bias task"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# dependencies
library(ez)
library(schoRsch)
library(tidyverse)
library(timesavers)
library(broom)
library(knitr)
library(kableExtra)

# knitr options
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

# get data
# shooter <- read.csv("shooter task data.csv", head = TRUE)  # original data
shooter <- read.csv("shooter_task_rawdata.csv", head = TRUE)  # original data
```

# Signal detection analysis

Code refactored from https://osf.io/rq6h2/. Original implementation used ddply, I refactored to tidyverse and checked for identical output.

## Process hit and false alarm data

```{r}
trimmed <- shooter %>%
  filter(values.result != "" &                 # exclude trials with no response
           values.result != "Noresponse") %>%    # exclude practice block trials that do not record a response type
  mutate(trialcode = as.character(trialcode),
         gun = ifelse(contains(trialcode, "Gun"), "gun", 
                      ifelse(contains(trialcode, "Object"), "object", NA)),
         Ethnicity = ifelse(contains("Black"), "Black", 
                            ifelse(contains("White"), "White", NA)))

# total trials per subject per trial type
temp1 <- trimmed %>%
  count(subject, Ethnicity, gun) %>%
  rename(totaltrials = n)

# n trials per subject per trial type per response type
temp2 <- trimmed %>%
  count(subject, Ethnicity, gun, values.result) %>%
  rename(ntrials = n)

shooter.bias <- 
  # join dfs
  left_join(temp1, temp2, by = c("subject", "Ethnicity", "gun")) %>%
  # calculate percentages
  mutate(percentage = ntrials/totaltrials) %>%
  # reshape
  spread(values.result, percentage) %>%
  rename(percentage.Hit  = Hit, 
         percentage.Miss = Miss, 
         percentage.FA   = FA, 
         percentage.CR   = CR) %>%
  mutate(
    # add zeros where data NAs are present
    percentage.Hit   = ifelse(is.na(percentage.Hit), 0, percentage.Hit),
    percentage.Miss  = ifelse(is.na(percentage.Miss), 0, percentage.Miss),
    percentage.FA    = ifelse(is.na(percentage.FA), 0, percentage.FA),
    percentage.CR    = ifelse(is.na(percentage.CR), 0, percentage.CR)
  ) %>%
  
  group_by(subject, Ethnicity, gun) %>%
  summarize(totaltrials = sum(ntrials, na.rm = TRUE),
            percentage.Hit  = sum(percentage.Hit, na.rm = TRUE), 
            percentage.Miss = sum(percentage.Miss, na.rm = TRUE), 
            percentage.FA   = sum(percentage.FA, na.rm = TRUE), 
            percentage.CR   = sum(percentage.CR, na.rm = TRUE)) %>%
  
  rowwise() %>%
  mutate(
    # invert Hits
    percentage.Hit = ifelse(gun == "Gun" & percentage.Hit == 1,
                            (1-(1/(2*totaltrials))),
                            percentage.Hit),
    
    # invert FAs
    percentage.FA = ifelse(gun == "Object" & percentage.FA == 0,
                           (1/(2*totaltrials)),
                           percentage.FA)
  ) %>%
  ungroup() %>%
  
  # summarize across trial types
  group_by(subject, Ethnicity) %>%
  summarise(percentage.Hit = sum(percentage.Hit), 
            percentage.FA = sum(percentage.FA)) %>%
  
  # compute d' and c
  mutate(d = round(qnorm(percentage.Hit) - qnorm(percentage.FA), 3),
         c = round(-0.5*(qnorm(percentage.Hit) + qnorm(percentage.FA)), 3),
         percentage.Hit = round(percentage.Hit, 3),
         percentage.FA  = round(percentage.FA, 3))
```

## Analyses

### d' between conditions

```{r}
attach(shooter.bias)

t.test(d ~ Ethnicity, paired = TRUE) %>%
  tidy()
```

### c between conditions

```{r}
t.test(c ~ Ethnicity, paired = TRUE) %>%
  tidy()
```

### Descriptives

```{r}
shooter.bias %>%
  group_by(Ethnicity) %>%
  summarize(mean.d = mean(d), 
            sd.d = sd(d),
            mean.c = mean(c), 
            sd.c = sd(c), 
            d.lower.ci = mean.d - (1.96*sd.d), 
            d.upper.ci = mean.d + (1.96*sd.d), 
            c.lower.ci = mean.c - (1.96*sd.c), 
            c.upper.ci = mean.c + (1.96*sd.c), 
            n = length(subject)) %>%
  round_df(2) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

NB RT and error analyses are also possible, but these are comparatively more common, and easy to implement in lm()/lmer().

